\framework: Morphology-based Entity and Relational Entity Extraction Framework for Arabic

### 2018-04-20 tex/merf/TALMajor/answer/Reviewer3.comments.tex
done 25 pages

@@

\answer{
We provide running examples with the definitions....
}

@@

\answer{
Our evaluation is a survey based evaluation.
Users used \framework to define their \framework based annotators. 
Users also built finite state transducer based annotators to 
perform the same tasks on the same corpora. 
We measured and reported the time needed by both.

We also performed manual annotation to construct the correctness 
benchmark.
We did that by correcting the automated annotations of both annotators
for the benchmark corpora. 
That alone took much more time than the time it took to develop and run
the \framework annotators.

We clarify that in the beginning of the Results 
Section ~\ref{sec:results}.
}

\itodo{
  Ameen: lets add the above to the section. 
}

@@

\answer{
  We include more information about the data and 
  the evaluation process.
  We include the size of the annotated data sets that 
  we used for accuracy evaluation in Table~\ref{tab:results}. 
  We discuss how the golden reference was constructed. 
  We also discuss the workload of the research assistant who built
  the task specific case studies. 
}

@@

\answer{
We extend the discussion section to discuss ... 
}

>>>>

\answer{
  We will provide a GUI tutorial file on the tool's website. 
}

@@

\answer{
We now provide running examples for the definitions in the 
MEF simulator Section. 
}

@@

\answer{

We address the comment by adding the following at the beginning
of the Results Section~\ref{sec:results}.

\begin{quote}
In this section we evaluate \framework with four case studies. 
We perform a survey like evaluation where developers manually 
built task specific information extraction tools for the case studies 
and other developers built equivalent \framework tools. 
The aim of the comparison is to showcase that \framework enables 
fast development of linguistic applications with similar accuracy 
and a reasonable affordable overhead in computational time. 
We report development time,
size of developed code versus size of grammar,
running time, and 
precision-recall as metrics
of cost, complexity, overhead, and accuracy,
respectively. 
\end{quote}

Developers implemented the same tasks and ran on the same corpora. 
%
We also performed manual annotation to construct the golden reference. 
We did that by correcting the automated annotations of both annotators
for the benchmark corpora. 
We clarify that in the Results Section~\ref{sec:results}.
}

@@

\answer{
  We address this in the Results Section~\ref{sec:results}
  as follows. 
  \begin{itemize} 
    \item 
  We include more information about the data and 
  the evaluation process.
\item 
  We include the size of the annotated data sets that 
  we used for accuracy evaluation in Table~\ref{tab:results}. 
\item 
  We now discuss how the golden reference was constructed. 
\item 
  We now discuss how precision and recall are computed as ratios of 
  entities correctly detected against extracted entities, and 
  entities correctly detected against the total number of entities, 
  respectively. 
\item 
  We also discuss the workload of the research assistant who built
  the task specific case studies. 
  \end{itemize} 
}

@@

\answer{
  We added  the following to the discussion Section. 
  \begin{quote}
We notice that ANGE, ATEEMA, and Genealogy tree report higher 
    precision than \framework. 
This is mainly due to their capacity 
to learn words and relations that may not have a match in the 
morphological analyzer based on co-occurence relations. 
For example, the sequence $p_1 t_1 p_2$ where $p_1$ and 
$p_2$ are persons and $t_1$ is a tell relationship helps
indicate that $x$ is a tell relationship in $p_1 x p_2$ 
even if the morphological analyzer did not return the required
feature for $x$ to match a tell relationship. 
\framework does not have that capacity yet unless it is
encoded in the C++ actions. 
  \end{quote} 
}

### 2018-04-20 tex/merf/TALMajor/answer/Reviewer4.comments.tex
done 25 pages

\item \label{Review.4.3} 
In the results section, there is no standard evaluation. 
The evaluation is based on comparisons with non-standard 
task-specific tools. 
Also, the reported precision and recall are very sensitive 
to the quality of the annotator!

\answer{
  Same answer to reviewer 3 results... 
}

>>>>

\item \label{Review.4.3} 
In the results section, there is no standard evaluation. 
The evaluation is based on comparisons with non-standard 
task-specific tools.

\answer{
  We modified the Results Section~\ref{sec:results} to 
    clarify our evaluation method as follows.
    \begin{itemize}
      \item We compare two different approaches trying to achieve the 
        same results using the same input corpora and the same
        golden reference. 
      \item We use standard metrics to report the results: development
        time, size of solution, runtime, precision and recall.
      \item We use the reported task-specific implementations because, 
        up to our knowledge, no other accessible relational entity 
        extraction tools exist for Arabic. 
    \end{itemize} 
}

\item \label{Review.4.4} 
Also, the reported precision and recall are very sensitive 
to the quality of the annotator!

\answer{
  This is correct. We added a precision discussion paragraph
    to reflect that in the Discussion Section~\ref{sec:discuss}. 
}

### 2018-04-20 tex/merf/TALMajor/discussion.tex
done till 3.19
done 25 pages

The two levels of interaction allow the user to separate between concepts 
that relate to word features, and more sophisticated entities 
that relate to sequences and context. 
%
The MBF, MRE, and user defined relations 
can be used to generate large annotated corpora in a fast manner. 
\framework visualization can be used later to edit the corpora 
and fix the annotations.

>>>>

The two levels of interaction allow the user to separate between concepts 
that relate to word features, and more sophisticated entities 
that relate to sequences and context. 
%
The MBF, MRE, and user defined relations 
can be used to generate large annotated corpora in a fast manner. 
\framework visualization can be used later to edit the corpora 
and fix the annotations.
%
\major{
The case studies showed that 
\framework requires some linguistic expertise to successfully execute the 
tasks. 
In contrast, the case specific implementations require more sophisticated 
linguistic and programming expertise to attain similar results.

We notice that ANGE, ATEEMA, and Genealogy tree report higher precision than \framework. 
This is mainly due to their capacity 
to learn words and relations that may not have a match in the 
morphological analyzer based on co-occurence relations. 
For example, the sequence $p_1 t_1 p_2$ where $p_1$ and 
$p_2$ are persons and $t_1$ is a tell relationship helps
indicate that $x$ is a tell relationship in $p_1 x p_2$ 
even if the morphological analyzer did not return the required
feature for $x$ to match a tell relationship. 
\framework does not have that capacity yet unless it is
encoded in the C++ actions. 
}

### 2018-04-20 tex/merf/TALMajor/introduction.tex
left with 3.3 3.20 3.24 3.27 and 4.3
done 25 pages

This work significantly extends the position paper~\cite{JaZaMatar} 
that allows for manual, and basic morphology annotation.
\framework enables a regular user to incrementally create 
complex annotations for Arabic text based on automatic 
extraction of morphological tags through a user friendly interactive interface. 
\framework has the following advantages.
\begin{itemize}
  \item \framework provides a novel and intuitive visual interface to build formulae over morphological features, 
    build regular expressions over the resulting formulae, and thereafter compute automatic tags.
  \item To our knowledge, this morphology-based framework is the first for Arabic entity and relational entity extraction.
  \item \framework provides the user with the ability to rapidly create annotated Arabic text corpora with sophisticated morphology-based tags.
\end{itemize}

@@

The rest of the paper is structured as follows.
%In the rest of this paper,
%Section~\ref{sec:motivation} presents a motivational example. 
Section~\ref{sec:morph} introduces Arabic morphological analysis 
and its important role in Arabic NLP.
%We present preliminaries necessary to understand the work 
%in Section~\ref{sec:preliminaries}. 
Section~\ref{sec:framework} presents the \framework components.
Section~\ref{sec:gui} presents the visual and user friendly interface of \framework. 
Section~\ref{sec:methodology} explains the methodology used to design \framework.
Section~\ref{sec:related} presents and discusses related work.
Section~\ref{sec:results} presents the evaluation results. 
Finally, we conclude and discuss future work in Section~\ref{sec:conclusion}.

>>>>

This work significantly extends~\cite{JaZaMatar} 
that allows for manual, and morphology annotation.
\framework enables a user to incrementally create 
complex annotations for Arabic based on automatic 
extraction of morphological tags through a user friendly interactive interface. 
\framework has the following advantages.
\begin{itemize}
  \item \framework provides a novel and intuitive visual interface to build formulae over morphological features, 
    build regular expressions over the resulting formulae, and thereafter compute automatic tags.
  \item To our knowledge, this morphology-based framework is the first for Arabic entity and relational entity extraction.
  \item \framework provides the user with the ability to rapidly create annotated Arabic text corpora with sophisticated morphology-based tags.
\end{itemize}

@@

The rest of the paper is structured as follows.
%In the rest of this paper,
%Section~\ref{sec:motivation} presents a motivational example. 
Section~\ref{sec:morph} introduces Arabic morphological analysis 
and its important role in Arabic NLP.
%We present preliminaries necessary to understand the work 
%in Section~\ref{sec:preliminaries}. 
Section~\ref{sec:methodology} explains the methodology of \framework.
Section~\ref{sec:framework} presents \framework components.
Section~\ref{sec:gui} presents \framework GUI. 
Section~\ref{sec:related} presents and discusses related work.
Section~\ref{sec:results} presents the evaluation results. 
Finally, we conclude and discuss future work in Section~\ref{sec:conclusion}.

### 2018-04-20 tex/merf/TALMajor/methodology.tex
done 25 pages

Figure~\ref{f:overview} illustrates the four processes involved in 
\framework methodology. 
The first process takes the Arabic text input and provides the user 
with a morphology-based Boolean (MB) formulae GUI. 
The user interactively composes MB-formulae using the GUI and the output 
of the simulator and the $Syn^k$ detector. 
The simulator and the detector apply the formulae over the morphological solutions 
of the Arabic text and produce the MB-formulae tags.

@@

After the relation construction phase, we are interested to relate the 
discovered entities that express the same concept.
\framework provides the extended synonym feature of second order as 
a default cross-reference relation ($Syn^2$).
In Figure~\ref{fig:intromotiv}, triggering this feature creates the 
edge labels with \cci{isSyn} 
between the nodes \cci{Khalifa Tower} and \cci{The building}.

@@

The user may inspect and refine the defined formulae, rules and relations and 
the corresponding formulae tags, rule tags, entities and relational entities
either using the GUI or directly through readable output files. 
The files are in the javascript object notation (JSON)~\cite{nolan2014javascript} 
format that is intuitive to read and modify. 
\framework separates the user defined formulae, rules, actions and relations in
a \framework tag type file and the 
matching tags in a tags files. 
The separation serves the user to apply the tag types to multiple case studies 
and to obtain a separate resulting tags file for each case study.

>>>>

Figure~\ref{f:overview} illustrates the four processes involved in 
\framework methodology. 
The first process takes Arabic text and provides the user 
with a morphology-based Boolean (MB) formulae GUI. 
The user interactively composes MB-formulae using the GUI and the output 
of the simulator and the $Syn^k$ detector. 
The simulator and the detector apply the formulae over the morphological solutions 
of the Arabic text and produce the MB-formulae tags.

@@

After relation extraction, we are interested to relate 
entities that express the same concept.
\framework provides the extended synonym feature of second order as 
a default cross-reference relation ($Syn^2$).
In Figure~\ref{fig:intromotiv}, triggering this feature creates the 
edge labels with \cci{isSyn} 
between the nodes \cci{Khalifa Tower} and \cci{The building}.

@@

The user may inspect and refine the defined formulae, rules and relations and 
the corresponding formulae tags, rule tags, entities and relational entities
either using the GUI or directly through readable output files. 
The files are in the javascript object notation (JSON)~\cite{nolan2014javascript} 
format that is intuitive to read and modify. 
\framework separates the user defined formulae, rules, actions and relations in
a \framework tag type file and the 
matching tags in a tags files. 
The separation serves the user to apply the tag types to multiple case studies 
and to obtain a separate file of resulting tags for each.

### 2018-04-20 tex/merf/TALMajor/results.tex
done 25 pages

@@

In this section we compare \framework implementations of 
the narrator chain, temporal entity, and genealogy entity extraction 
tasks to the task specific techniques proposed to solve them
in ANGE~\cite{ZaMaFlairs2012HadithBio}, 
ATEEMA~\cite{ZaMa2012IJCLATime},  and
GENTREE~\cite{ZaMaHaCicling2012Entity}, respectively. 
We also compare a \framework number normalization task to 
a task specific implementation. 
We evaluated ANGE with
Musnad Ahmad, a hadith book, where we constructed an annotated 
golden reference containing 1,865 words.
We evaluated ATEEMA 
with articles from issues of  the Lebanese 
Al-Akhbar newspaper where we constructed 
an annotated golden reference containing 1,677 words. 
For the genealogical tree extraction we used 
an extract from the Genesis biblical text with 1,227 words.
Finally, we used an annotated article from the Lebanese Assafir
newspaper with 1,399 words to evaluate the NUMNORM case study
\footnote{available at \url{http://www.assafir.com} and 
\url{http://www.al-akhbar.com}.}. 
%
In the online appendix%
~\footnote{available at ~\url{-ommited-for-anonymity-}}%http://webfea.fea.aub.edu.lb/fadi/pdfs/merfappendix.pdf
, we report on eight additional \framework case studies.
\major{
Manual annotators inspected the outcome and provided corrections where tools 
made mistakes.
The corrections form the manual gold annotation that we compared against.
}

Table~\ref{tab:results} reports the development time,
extraction runtime, recall and precision 
of the output MRE tags, 
the size of the task in lines of code or in number of \framework rules, 
for both the standalone task specific and the \framework implementations.
\major{
The development time measures the time required for 
developing the case study.
For instance, ANGE~\cite{ZaMaFlairs2012HadithBio} required two 
months of development by a research assistant with six and 14 hours of 
course work and teaching duties, respectively.
}

@@

\setcode{utf8}
\setarab
\begin{table}[tb!]
  \centering
  \caption{Narrator chain example.}
  \begin{Verbatim}[xleftmargin=1.5cm,fontsize=\relsize{-1},commandchars=\\\{\},codes={\catcode`$=3 \catcode`_=8}]
name:   PN ((MEAN)? PN)*;
nar:    name ((NONE)^3 FAM (NONE)^3 name)*;
pbuh:   BLESS GOD UPONHIM GREET;
nchain: ($s_1=$TOLD $s_2=$nar)+ ((PN|FAM|NONE)^8 pbuh)?
\end{Verbatim}
\resizebox{0.8\columnwidth}{!}{
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
    \toprule 
    \notrrl{القعقاع} & \notrrl{بن} & \notrrl{عمارة} & \notrrl{عن} & \notrrl{جرير} & \notrrl{حدثنا} & \notrrl{سعيد} & \notrrl{بن} & \notrrl{قتيبة} & \notrrl{حدثنا} \\
    \midrule 
    \noarrl{القعقاع} & \noarrl{بن} & \noarrl{عمارة} & \noarrl{عن} & \noarrl{جرير} & \noarrl{حدثنا} & \noarrl{سعيد} & \noarrl{بن} & \noarrl{قتيبة} & \noarrl{حدثنا} \\
    \midrule
    PN    & FAM   & PN    & TOLD  & PN    & TOLD  & PN    & FAM   & PN    & TOLD \\
    \midrule
    name & & name & & name & & name &  & name & \\
    \midrule
    \multicolumn{3}{|c|}{nar} &       & nar   &       & \multicolumn{3}{c|}{nar} &  \\
    \midrule
    \multicolumn{10}{|c|}{nchain}
    \\
    \bottomrule
    \end{tabular}%
}
  \label{tab:nchain}%
\end{table}%
\setcode{standard}

@@

For brevity, we omit the details of \framework temporal entity extraction, 
genealogy tree, and number normalization case studies, describe them shortly
below and provide a full description in the online Appendix.

\begin{table}[tb!]
  \centering
  \caption{\framework MBF and user-defined relation accuracy }
  \resizebox{0.65\columnwidth}{!}{
    \begin{tabular}{l|c|c|c|c}
     \toprule
     \multirow{2}{*}{Task} & \multicolumn{2}{c|}{MBF accuracy} & \multicolumn{2}{c}{relation accuracy}\\
     & Recall & Precision & Recall & Precision \\
    \midrule
    Narrator chain & 0.99 & 0.85 & 0.99 & 0.98 \\
    Number normalization & 0.99 & 0.99 & 0.97 & 0.95 \\
    Temporal entity & 0.99 & 0.52 & 0.98 & 0.89 \\
    Genealogy tree & 0.99 & 0.75 & 0.81 & 0.96 \\
    \bottomrule
    \end{tabular}%
    }
  \label{tab:mbfer}%
\end{table}%

\vspace{-1em}
\subsection{Temporal entity extraction}

Temporal entities are text chunks that express temporal information. 
Some represent absolute time such as \RL{Al_hAms mn 'Ab 2010}. 
Others represent relative time such as \RL{b`d _hmsT 'ayAm}, and quantities 
such as \RL{14 ywmA}. 
{\em ATEEMA} presents a temporal entity detection technique for the Arabic language using 
morphological analysis and finite state transducers~\cite{ZaMa2012IJCLATime}.

Table~\ref{tab:mbfer} shows that \framework detected almost all the MBF matches with 99\% recall, 
however it shows low precision (52\%). 
As for the semantic relation construction, \framework presents a 98\% recall and 89\% precision.

\vspace{-1em}
\subsection{Genealogy tree}

Biblical genealogical lists trace key biblical figures such as Israelite kings and
prophets with family relations. 
The family relations include wife and parenthood. 
A sample genealogical chunk of text is \RL{w wld hArAn lw.tA} 
meaning ``and Haran became the father of Lot''.

GENTREE~\cite{ZaMaHaCicling2012Entity} 
automatically extracts the genealogical family trees using morphology, 
finite state machines, and graph transformations. 
Table~\ref{tab:mbfer} shows that \framework detected 
MBF matches with 99\% recall, and 75\% precision, and
extracted user-defined relations with 81\% recall and 96\% precision.

\newcommand*{\fvtextcolor}[2]{\textcolor{#1}{#2}}
\begin{figure}[tb!]
\centering
  \begin{tabular}{p{5.5cm}p{5.5cm}}
\begin{Verbatim}[fontsize=\relsize{-1},frame=single,label=TMB algorithm,commandchars=\\\[\]] 
cout << \fvtextcolor[red][$s1.text];
if(isHundred) {
  if(current != 0) {
    previous += current;
  }
  current = currentH * \fvtextcolor[red][$s1.number];
  currentH = 0;
  isHundred = false;
  isKey = true;
} else if(current == 0) {

  current = \fvtextcolor[red][$s1.number];
  isKey = true;
} else if(!isKey) {

  isKey = true;
  current = current * \fvtextcolor[red][$s1.number];
} else {
  previous += current;
  current = \fvtextcolor[red][$s1.number];
}
\end{Verbatim}
&
\begin{Verbatim}[fontsize=\relsize{-1},frame=single,label=DT algorithm,commandchars=\\\[\]] 
if(isHundred) {
  currentH += \fvtextcolor[red][$s0.number];
} else if(current == 0) {
  current = \fvtextcolor[red][$s0.number];
} else if(isKey) {
  previous += current;
  current = \fvtextcolor[red][$s0.number];
} else {
  current += \fvtextcolor[red][$s0.number]; }
isKey = false;
\end{Verbatim}
\begin{Verbatim}[fontsize=\relsize{-1},frame=single,label=H algorithm,commandchars=\\\[\]] 
isHundred = true;
if(current == 0)  {
  currentH = \fvtextcolor[red][$s2.number];
} else if(!isKey) {
  currentH = current * \fvtextcolor[red][$s2.number];
  current = 0;
} else {
  currentH = \fvtextcolor[red][$s2.number];}
isKey = false;
\end{Verbatim}
\\ 
\end{tabular}
\caption{Actions for TMB, DT, and H MRE expressions.}
\label{fig:numnormalgo}
\end{figure}

@@

We implemented a number normalization extractor using \framework and 
compared it with {\em NUMNORM}, a 
C++ implementation for number normalization. 
First, we defined the MBFs \cci{DT}, \cci{H}, and \cci{TMB}
to denote (1) digits and tens, (2) hundreds, and (3) 
thousands, millions, and billions, respectively.
The \cci{num} MRE 
\cci{(DT|TMB|H)+} is one or more \cci{DT}, \cci{TMB}, or \cci{H} tags. 
\framework assigns the symbols $s_1$, $s_2$, and $s_3$ 
for the sub-expressions \cci{DT}, \cci{TMB}, and \cci{H}, respectively. 
%The actions associated with the sub-expressions \cci{DT}, \cci{TMB}, and \cci{H} are presented in the appendix.
\major{
Figure~\ref{fig:numnormalgo} shows the actions associated with the \cci{DT}, \cci{TMB}, and \cci{H} subexpressions that cumulatively compute the numeric value of the numeric expression match.
The actions use \framework API to access features of the matches such 
as text (\cci{\$s1.text}) and numeric 
value (\cci{\$s1.number}) of literal numbers such as numbers from one to ten.
}
%
Table~\ref{tab:mbfer} shows high accuracy in MBF tagging with 99\% recall and 99\% precision, and
high accuracy in relation extraction with 97\% recall and 95\% precision.

\vspace{-1em}
\subsection{Discussion}
\label{subsec:discuss}
\label{sec:discuss}
\input{discussion}

>>>>

In this section we evaluate \framework with four case studies. 
We perform a survey like evaluation where developers manually 
built task specific information extraction tools for the case studies 
and other developers built equivalent \framework tools. 
The aim of the comparison is to showcase that \framework enables 
fast development of linguistic applications with similar accuracy 
and a reasonable affordable overhead in computational time. 
We report development time,
size of developed code versus size of grammar,
running time, and 
precision-recall as metrics
of cost, complexity, overhead, and accuracy,
respectively.

We survey three case studies from the literature: 
(1) narrator chain, (2) temporal entity, and (3) genealogy entity extraction 
tasks and we use the reported development time for the task specific 
techniques proposed in
ANGE~\cite{ZaMaFlairs2012HadithBio}, 
ATEEMA~\cite{ZaMa2012IJCLATime},  and
GENTREE~\cite{ZaMaHaCicling2012Entity}, respectively. 
%
We also compare a \framework number normalization task to 
a task specific implementation.

We evaluated ANGE with
Musnad Ahmad, a hadith book, where we constructed an annotated 
golden reference containing 1,865 words.
We evaluated ATEEMA 
with articles from issues of  the Lebanese 
Al-Akhbar newspaper where we constructed 
an annotated golden reference containing 1,677 words. 
For the genealogical tree extraction we used 
an extract from the Genesis biblical text with 1,227 words.
Finally, we used an annotated article from the Lebanese Assafir
newspaper with 1,399 words to evaluate the NUMNORM case study
\footnote{available at \url{http://www.assafir.com} and 
\url{http://www.al-akhbar.com}.}. 
%
In the online appendix%
~\footnote{available at ~\url{-ommited-for-anonymity-}}%http://webfea.fea.aub.edu.lb/fadi/pdfs/merfappendix.pdf
, we report on eight additional \framework case studies.
\major{
Manual annotators inspected the outcome and provided corrections where tools 
made mistakes.
The corrections form the manual gold annotation that we compared against.
}

@@

Table~\ref{tab:results} reports the development time,
extraction runtime, recall and precision 
of the output MRE tags, 
the size of the task in lines of code or in number of \framework rules, 
for both the standalone task specific and the \framework implementations.
\major{
The development time measures the time required for 
developing the case study.
For instance, ANGE~\cite{ZaMaFlairs2012HadithBio} required two 
months of development by a research assistant with six and 14 hours of 
course work and teaching duties, respectively.
Recall refers to the fraction of the entities correctly detected against the
total number of entities. 
Precision refers to the fraction of correctly
detected entities against the total number of extracted entities. 
}

@@

\setcode{utf8}
\setarab
\begin{table}[tb!]
  \centering
  \caption{Narrator chain example.}
  \begin{Verbatim}[xleftmargin=1.5cm,fontsize=\relsize{-1},commandchars=\\\{\},codes={\catcode`$=3 \catcode`_=8}]
name:   PN ((MEAN)? PN)*;
nar:    name ((NONE)^3 FAM (NONE)^3 name)*;
pbuh:   BLESS GOD UPONHIM GREET;
nchain: ($s_1=$TOLD $s_2=$nar)+ ((PN|FAM|NONE)^8 pbuh)?
\end{Verbatim}
\resizebox{0.8\columnwidth}{!}{
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
    \toprule 
    \notrrl{القعقاع} & \notrrl{بن} & \notrrl{عمارة} & \notrrl{عن} & \notrrl{جرير} & \notrrl{حدثنا} & \notrrl{سعيد} & \notrrl{بن} & \notrrl{قتيبة} & \notrrl{حدثنا} \\
    \midrule 
    \noarrl{القعقاع} & \noarrl{بن} & \noarrl{عمارة} & \noarrl{عن} & \noarrl{جرير} & \noarrl{حدثنا} & \noarrl{سعيد} & \noarrl{بن} & \noarrl{قتيبة} & \noarrl{حدثنا} \\
    \midrule
    PN    & FAM   & PN    & TOLD  & PN    & TOLD  & PN    & FAM   & PN    & TOLD \\
    \midrule
    name & & name & & name & & name &  & name & \\
    \midrule
    \multicolumn{3}{|c|}{nar} &       & nar   &       & \multicolumn{3}{c|}{nar} &  \\
    \midrule
    \multicolumn{10}{|c|}{nchain}
    \\
    \bottomrule
    \end{tabular}%
}
  \label{tab:nchain}%
  \vspace{-1.5em}
\end{table}%
\setcode{standard}

@@

%For brevity, we omit the details of \framework temporal entity extraction, 
%genealogy tree, and number normalization case studies, describe them shortly
%below and provide a full description in the online Appendix.

\begin{table}[tb!]
  \centering
  \caption{\framework MBF and user-defined relation accuracy }
  \resizebox{0.65\columnwidth}{!}{
    \begin{tabular}{l|c|c|c|c}
     \toprule
     \multirow{2}{*}{Task} & \multicolumn{2}{c|}{MBF accuracy} & \multicolumn{2}{c}{relation accuracy}\\
     & Recall & Precision & Recall & Precision \\
    \midrule
    Narrator chain & 0.99 & 0.85 & 0.99 & 0.98 \\
    Number normalization & 0.99 & 0.99 & 0.97 & 0.95 \\
    Temporal entity & 0.99 & 0.52 & 0.98 & 0.89 \\
    Genealogy tree & 0.99 & 0.75 & 0.81 & 0.96 \\
    \bottomrule
    \end{tabular}%
    }
  \label{tab:mbfer}%
  \vspace{-1.5em}
\end{table}%

\vspace{-1.5em}
\subsection{Temporal entity extraction}
\vspace{-1em}

Temporal entities are text chunks that express temporal information. 
Some represent absolute time such as \RL{Al_hAms mn 'Ab 2010}. 
Others represent relative time such as \RL{b`d _hmsT 'ayAm}, and quantities 
such as \RL{14 ywmA}. 
{\em ATEEMA} presents a temporal entity detection technique for the Arabic language using 
morphological analysis and finite state transducers~\cite{ZaMa2012IJCLATime}. 
%
Table~\ref{tab:mbfer} shows that \framework detected almost all the MBF matches with 99\% recall, 
however it shows low precision (52\%). 
As for the semantic relation construction, \framework presents a 98\% recall and 89\% precision.

\vspace{-2em}
\subsection{Genealogy tree}
\vspace{-1em}

Biblical genealogical lists trace key biblical figures such as Israelite kings and
prophets with family relations. 
The family relations include wife and parenthood. 
A sample genealogical chunk of text is \RL{w wld hArAn lw.tA} 
meaning ``and Haran became the father of Lot''.
%
GENTREE~\cite{ZaMaHaCicling2012Entity} 
automatically extracts the genealogical family trees using morphology, 
finite state machines, and graph transformations. 
Table~\ref{tab:mbfer} shows that \framework detected 
MBF matches with 99\% recall, and 75\% precision, and
extracted relations with 81\% recall and 96\% precision.

\newcommand*{\fvtextcolor}[2]{\textcolor{#1}{#2}}
\begin{figure}[tb!]
\centering
  \begin{tabular}{p{5.5cm}p{5.8cm}}
\begin{Verbatim}[fontsize=\relsize{-1},frame=single,label=TMB algorithm,commandchars=\\\[\]] 
cout << \fvtextcolor[red][$s1.text];
if(isHundred) {
  if(current != 0) {
    previous += current;
  }
  current = currentH * \fvtextcolor[red][$s1.number];
  currentH = 0;
  isHundred = false;
  isKey = true;
} else if(current == 0) {
  current = \fvtextcolor[red][$s1.number];
  isKey = true;
} else if(!isKey) {
  isKey = true;
  current = current * \fvtextcolor[red][$s1.number];
} else {
  previous += current;
  current = \fvtextcolor[red][$s1.number];}
\end{Verbatim}
&
\begin{Verbatim}[fontsize=\relsize{-1},frame=single,label=DT algorithm,commandchars=\\\[\]] 
if(isHundred) {currentH += \fvtextcolor[red][$s0.number];
} else if(current == 0) {
  current = \fvtextcolor[red][$s0.number];
} else if(isKey) {
  previous += current;
  current = \fvtextcolor[red][$s0.number];
} else {current += \fvtextcolor[red][$s0.number]; }
isKey = false;
\end{Verbatim}
\begin{Verbatim}[fontsize=\relsize{-1},frame=single,label=H algorithm,commandchars=\\\[\]] 
isHundred = true;
if(current == 0)  {
  currentH = \fvtextcolor[red][$s2.number];
} else if(!isKey) {
  currentH = current * \fvtextcolor[red][$s2.number];
  current = 0;
} else {currentH = \fvtextcolor[red][$s2.number];}
isKey = false;
\end{Verbatim}
\\ 
\end{tabular}
  \vspace{-2em}
\caption{Actions for TMB, DT, and H MRE expressions.}
  \vspace{-1.5em}
\label{fig:numnormalgo}
\end{figure}

@@

We implemented a number normalization extractor using \framework and 
compared it with {\em NUMNORM}, a 
C++ implementation for number normalization. 
First, we defined the MBFs \cci{DT}, \cci{H}, and \cci{TMB}
to denote (1) digits and tens, (2) hundreds, and (3) 
thousands, millions, and billions, respectively.
The \cci{num} MRE 
\cci{(DT|TMB|H)+} is one or more \cci{DT}, \cci{TMB}, or \cci{H} tags. 
\framework assigns the symbols $s_1$, $s_2$, and $s_3$ 
for the sub-expressions \cci{DT}, \cci{TMB}, and \cci{H}, respectively. 
%The actions associated with the sub-expressions \cci{DT}, \cci{TMB}, and \cci{H} are presented in the appendix.
\major{
Figure~\ref{fig:numnormalgo} shows the actions associated with the \cci{DT}, \cci{TMB}, and \cci{H} subexpressions that cumulatively compute the numeric value of the numeric expression match.
The actions use \framework API to access features of the matches such 
as text (\cci{\$s1.text}) and numeric 
value (\cci{\$s1.number}) of literal numbers such as numbers from one to ten.
}
%
Table~\ref{tab:mbfer} shows high accuracy in MBF tagging and relation extraction 
with 99\% and 97\% recall and 99\% and 95\% precision, respectively.

\vspace{-1.5em}
\subsection{Discussion}
\label{subsec:discuss}
\label{sec:discuss}
\input{discussion}
